Sperm do not evolve to collaborate in female meiotic drive.

### 2014-03-04 writeup_and_notes/male_dependent_drive.tex
minor typo fixing and wordsmithing, also ref. updating

%\title{Why does female meiosis allow the opportunity for exploitation by selfish sperm? BETTER TITLE?}
\title{Co-operation between the sexes trumps selfish control of female
  meiotic drive. Or 
Sperm do not evolve to collaborate in female meiotic drive. Or Why does female meiosis allow the opportunity for exploitation by selfish sperm? }

@@

Despite the apparent unity of the organism, alleles can
gain an evolutionary advantage at a cost to  individual fitness
\citep{Burt2006}, often by exploiting meiosis and gametogenesis.
Because only one of the four products of female meiosis is transmitted to the egg, female meiosis is particularly vulnerable to such exploitation \citep{Sandler1957,Pardo-ManuelDeVillena2001a}. 
An allele that biases female meiosis in its favor (i.e. a meiotic driver), may increase in frequency even if this driver entails a pleiotropic fitness cost \citep{Prout1973}, generating a genetic conflict between the success of the driver and organismal fitness.
Meiotic drivers observed in nature (in both plants
\citep{Buckler1999,Fishman2005,Fishman2008}, and animals
\citep{Agulnik1990,Wu2005,Pardo-ManuelDeVillena2001c}) highlight this
conflict -- the selfish benefits of drive and the associated
pleiotropic fitness costs sustain a balanced polymorphism
\citep{Prout1973}, 
and often generate an ongoing evolutionary escalation of drive suppressors and enhancers \citep{Dawe1996,Fishman2008}. 
The threat of meiotic drive to organismal fitness is potentially so
severe that many basic properties of
meiosis and o\"{o}genesis, including the initial genome doubling in
meiosis I \citep{Haig1991}, arrested female meiosis
\citep{Mira1998}, centromere machinery \citep{MalikHenikoff}, and sex differences in the recombination rate \citep{Haig2010,Brandvain2012} 
have perhaps evolved to disrupt meiotic drive and enforce fairness.

@@

It is certainly not the case that animals are mechanistically incapable of circumventing this requirement.
The stage at which female meiosis requires fertilization varies widely across taxa, and in a few animal taxa, 
	female meiosis is complete upon ovulation (that is fertilization is not required for female meiosis \citep[see
	Table 1 of ][]{Masui_book}).

It is also not the case that sperm is mechanistically incapable of influencing the outcome of female meiosis.
Mechanistic evidence for this possibility comes from
\emph{C. elegans} the fact that premature deployment of the aster (a vital component of mitotic
machinery) provided by the sperm can disrupting MII meiotic segregation
in the egg, leading to a triploid zygote \citep{McNally2012}. 
Additionally, genetic evidence suggests that the transmission patterns
in heterozygous females many potentially depend on sperm haplotype. 
Specifically, two female meiotic drive systems in mouse (In and Om), both operate by distorting the second meiotic division, 
and in both systems the outcome of female meiosis depends the genotype of the fertilizing sperm \citep{Agulnik1993,Wu2005}. 
While such genetic evidence is suggestive, we note the difficulty in
ruling out the alternative hypothesis of early selection on zygotes \citep[pages
52-54][]{Burt2006}.

@@

In the standard single-locus, biallelic model of female meiotic drive, the driving allele is transmitted to the egg  in heterozygotes with probability  $d > \frac{1}{2}$, regardless of sperm genotype. 
To depict a case of a self-promoting (or `green-bearded') meiotic
driver,  we modify this standard model such that the driver is only effective when fertilized by a sperm carrying that allele (see Figure
\ref{Eggsperm_2_allele_cartoon}) \footnote{the timing of fertilization relative to female meiosis places another constraint on $d$, for example, if fertilization (and therefore, sperm dependent drive) takes place at MII (as in mammals),
	female drive requires an uneven number of crossovers between the centromere and the drive locus, 
	so $d$ is bounded to be $<0.75$? \citep[see ][ for discussion]{Buckler1999}. Could also put this in
        caption of fig.}.
We then identify the conditions under which this self propagating driver can spread, and evaluate whether a driver of this form could generate a sustained conflict favoring the evolution of suppressors. 
\yb{EQUATIONS A1-A4}

We first consider a driving allele neutral in heterozygotes, 
	but deleterious in homozygotes (whose relative fitness is
        $1-s$). This model is laid out in eqn. \ref{eqn.A1}-\ref{eqn.A3}.
In this model, a standard meiotic driver can always invade because 
	when rare it occurs predominantly in heterozygotes and it therefore drives without a fitness cost. 
However, a severe homozygous fitness cost ($s>XXd$) prevents  
	drivers from fixing  \citep{Prout1973}, 
	providing an opportunity for the evolution of
	drive suppressors \citep{XX}, corresponding well to 
	empirical examples of female meiotic drive.

In contrast to a traditional driver, which drives but pays effectively
no fitness cost as a rare deleterious recessive, 
	a self-promoting driver creates homozygotes
        when it drives, and therefore there is a drive-associated fitness cost even when rare. 
Our self-promoting driver must overcome this homozygous fitness cost simply to spread when rare.
Therefore with a recessive fitness cost,
	the conditions allowing the invasion of a self-promoting driver
 	are far more restrictive than those for a standard meiotic driver (see Figure \ref{Invasion_space}).

@@

While the allelic identity of sperm could plausibly influence the outcome of female meiosis, 
	limited gene expression in sperm \citep[e.g.][]{Joseph2004}
	suggest a model where sperm influence female meiosis via expression of the fertilizing male's
	diploid genotype (perhaps due to proteins and RNAs packaged into the sperm).
This adds an additional parameter to the system -- the dominance of the drive enhancer in heterozygous males, 
	when this dominance is high, a narrow sliver of parameter space sustaining a polymorphism emerges (see Figure S\ref{Effect_of_dominance})
 presumably because positive frequency dependence diminishes with the dominance of the male effect driver.

@@

Above, we explored a model were a driver that drove in females when signaled by a complementary allele in sperm.  
We complement this single locus model of an allele with effects in
both male and with femalr, with  an alternative model of two linked loci - one female driver and the other a sperm allele which modifies the effect of drive upon fertilization. 
%Here, we examine an alternative scenario (and more realistic) in which
In this model initially a female meiotic driver, with no sperm-dependency, reaches drive-viability equilibrium (with two alleles
$A$ and $B$ are the ancestral non-driver and driver alleles).
Subsequently, a sperm-acting modifier of female meiotic drive arises tightly
        linked to this balanced locus (effectively creating a third
        allele/haplotype at this locus).
Assuming tight linkage between female driver and sperm modifier is consistent with the nature of well characterized meiotic drive systems 
	which are often maintained as polymorphic inversions with numerous linked modifiers (CITE).

@@

%%%sperm-acting facilitator
We first consider a sperm-acting facilitator of
	female meiotic drive tightly linked to the $B$ background, 
	generating a third allele $B^{+}$ (i.e. the non-recombining haplotype of a
        driver allele and sperm-acting drive modifier). 
This new allele acts in sperm to increase the effectiveness of both
	the $B$ and  $B^{+}$ drive in $AB$ and $AB^{+}$ heterozygotes, see Figure \ref{Eggsperm_3_allele_cartoon} 
	for a simple schematic of this drive model.  
Naively, the $B^{+}$ may spread by capitalizing on the additional drive it causes; hover,  
	this is not the case for a few simple reasons. 
First, the sperm-acting drive facilitator (the novel $B^{+}$ haplotype) 
	arises when the ancestral driver ($B$) is at drive-selection balance, 
	and therefore immediately suffers a homozygous fitness cost.  
Worse yet, a novel $B^{+}$ haplotype most often helps 
	the $B$  allele drive ($B^+$ sperm meeting $AB$ eggs), because $B$ is initially more common than $B^{+}$. 
As $B^{+}$ facilitates $B$ these interactions frequently form 
	$BB^{+}$ heterozygotes that suffer the same fitness costs as $BB$. 
Therefore, sperm-acting drive facilitator alleles experience a profound disadvantage 
	in this scenario, even more so than under the previous two allele model. 
We have found no parameter range of this
	three allele system that allows the sperm-acting drive facilitator $B^{+}$ to
	invade the population. \yb{MATHS?}

@@

Sexual reproduction is a high-stakes event that determines what gets transmitted to the next generation.  
As a consequence of this intense competition, alleles that gain a transmission advantage during reproduction 
	can succeed evolutionarily, even if they lower the 
        fitness of the population, 
	the mother whose offspring they will inhabit, 
	and even the organism in which they reside. 
This potential generates numerous conflicts  \cite{Burt2006} including sexual conflicts between mates \cite{Arnqvist2005}, 
	as well as conflicts between alleles that are overtransmitted in meiosis and the organisms they injure while doing so. 
Such conflicts and their resolution likely play a major role in the structure and evolution of many basic biological processes \citep{Rice2013}.

It seems that allowing sperm to influence the outcome of female meiosis would generate a confluence of these potential conflicts -- 
	sperm could actually assist an allele that distorts female meiosis.
However, this is not the case.
We find that an allele which acts through sperm to distort female meiosis in its favor %YB [too awkward / unclear?] 
	can rarely spread through a population if it bears any cost. 
Additionally, when this self-promoting driver can spread, it can only rarely 
	be maintained as a protected polymorphism, and due to its positive frequency dependence 
	it spends very little time at intermediate frequency.
As such, this type of exploitation cannot generate a sustained genetic conflict and so it is unlikely
	that female oogenesis and meiosis will evolve to prevent their effect.  
This likely explains why females can delay the completion of meiosis until after fertilization 
	without risking exploitation by collaborations by female
        drivers and sperm alleles.
%Male modification of recombination rates. \yb{trying to think through this \dots{}}

Why is it that an allele that biases female meiosis in its favor can generate a genetic conflict, but an allele in sperm that assists this female driver cannot? 
So long as the transmission advantage of female meiotic drive outweigh the organismal fitness cost to heterozygotes the female driver can spread when rare, and it increases in 		
	frequency until the fitness cost experienced when homozygous balances the transmission advantage.
By contrast a sperm promotor of female drive is only effective when matched with a heterozgyote female -- meaning that when rare this allele rarely enhances female drive. 
Even worse, when it does so it will preferentially find itself in a low fitness drive homozygote. 
Not only are drive-promoting sperm alleles unable to create a sustained genetic conflict, 
	but alleles in sperm with the opposite effect - that is those that prevent their own drive through female meiosis do maintain a polymorphisms and 
	provide evolution  with time and opportunity to further minimize drive.
This is because such drive suppressing alleles reduced their chances
of forming low fitness homozygotes. 
In fact, an allele in sperm which causes the alternate allele to drive can be favored (FIG), a counterintuitive result 
	consistent with the tenuous evidence that the non-driving
        $ln^-$ allele is over transmitted through female meiosis in $ln^-/ln^+$ heterozygote when fertilized by $ln^+$
        sperm \cite{agulnick, Pomiankowski1993}.   
More generally both linked and unlinked alleles may often spread that act through sperm to
reduce the opportunity of female meiotic drive in order to disable
segregating drive systems, by for example reducing the mechanistic
asymmetries that female drivers exploit.

@@

%Gametogenesis and meiosis are common times for genetic conflict. 
%This is a consequence of the fact that as the set of alleles that consitute
%	an individual are in the process of parting ways, 
%	with only a subset of them being assured transmission.
%In constrast, the perspective of the fertilizing sperm's genome is different, it has to live with the consequences of any distortion it causes 
%to female meiosis through the entire next life cycle of the organism. 
%It is therefore in the interest of the fertilizing sperm's genome to
%make a high fitness zygote even if that means alleles in sperm
%disrupting other copies of themselves from gaining a 
%transmission advantage through female meiosis.
  \appendix
\section*{Appendix}
In our first scenerio allele $A$ is the ancestral allele and allele
$B$ is our sperm dependent drive allele.
The sperm dependence of the drive is a form of assortative
transmission (mating), and so we have to keep track of the genotype frequencies.
Our genotype frequencies at birth are $x_{AA},~x_{AB},$ and
$x_{BB}$. These three genotypes have fitnesses $w_{AA},~w_{AB},$ and
$w_{BB}$, which are identical between males and females. As the
fitnesses are the same in males and females the frequencies of the
genotypes are identical in males and females, but in the following
equations they are marked with $\Mars$ and $\Venus$ respectively to
allow the reader to more easily follow the logic.
The frequency of the three genotypes at birth in the next generation is:
\begin{equation}
x'_{BB} =\frac{1}{\bar{w} } \Big( w_{BB} x_{BB}^{\Venus} (w_{BB} x_{BB}^{\Mars} +\frac{1}{2}
  w_{AB} x_{AB}^{\Mars} ) + w_{AB} x_{AB}^{\Venus} (d w_{BB}
  x_{BB}^{\Mars} +\frac{1}{2} d
w_{AB} x_{AB}^{\Mars} ) \Big),  \label{eqn.A1}
\end{equation}
\begin{equation}
x'_{AA} =\frac{1}{\bar{w} } \Big(  w_{AA} x_{AA}^{\Venus} (w_{AA} x_{AA}^{\Mars} + \frac{1}{2}
   w_{AB} x_{AB}^{\Mars}) + w_{AB} x_{AB}^{\Venus}(\frac{1}{2} w_{AA} x_{AA}^{\Mars}+ \frac{1}{2} 
  w_{AB} x_{AB}^{\Mars}) \Big),  \label{eqn.A2}
\end{equation}
and
\begin{eqnarray}
x'_{AB} = & \frac{1}{\bar{w} } \Big( w_{AA} x_{AA}^{\Venus}  ( w_{BB}  x_{BB}^{\Mars} + \frac{1}{2}
  w_{AB}  x_{AB}^{\Mars} ) +w_{BB} x_{BB}^{\Venus} ( w_{AA}  x_{AA}^{\Mars} + \frac{1}{2}
  w_{AB}  x_{AB}^{\Mars} )  \nonumber \\
& + x_{AB}^{\Venus}(\frac{1}{2}  w_{AA}  x_{AA}^{\Mars}+\frac{1}{2} (1-d)
 w_{AB}  x_{AB}^{\Mars} + \frac{1}{4}  w_{AB}
 x_{AB}^{\Mars} + (1-d)  w_{BB}  x_{BB}^{\Mars} ) \Big) \label{eqn.A3}
\end{eqnarray}
where the mean fitness of the population is
\begin{equation}
\bar{w} = w_{AA} x_{AA}+w_{AB}  x_{AB} +w_{BB}  x_{BB}.
\end{equation}
We iteration can then iterate this recursion over generations, for
example to produce the frequency trajectories shown in the inset
Figure \ref{Invasion_space} by plotting $x_{BB}+ \frac{1}{2}x_{AB}$
over time. We can also use this to assess the
invasion criteria for the allele by starting allele B at very low
frequency, and seeing if it increases across the generations.

\paragraph{cost of driver in heterozygotes}
We also explored a range of other costs to the drive alleles other than a
simple recessive cost, for example an additive cost. We modified the recursions given in eqn
\ref{eqn.A1}-\ref{eqn.A3} to feature an additive cost $s$ for each
copy of the $B$ allele. A none recessive cost to the drive allele can
cause bistability of the drive allele. The B allele is unable to enter
the population when rare, this results from the fact that
when the allele is rare ($x_B \ll 1$) sperm carrying it rarely
encounter heterozygous eggs ($\sim x^2$) but the allele still pays a
cost in heterozygous individuals.  While when the B allele is common ($x_B
\gg 0$) encounters between heterozygous eggs and $B$ allele sperm
become more common allowing drive to occur this can allow the $B$
allele to spread when common. In Supplementary Figure
\ref{bistable_additive} we show the unstable equilibrium frequency
above which the $B$ allele can invade. This equilibrium frequency was found by a simple
parameter search.  \gc{wonder if a bit of work could find this}

\paragraph{Dominance} In the supplement we also explore a model where
the genotype of the male who produced the fertilizing sperm affects female drive,
instead of just the allelic type of the sperm. We assume that if a
heterozygous egg ($AB$) is fertilized by a sperm with the following
male ($\Mars$) genotypes the $B$ allele will be transmitted through
female meiosis with the follow probabilites\\
\begin{tabular}{cccc}
$\Mars$ geno. & AA & AB  & BB\\
P(B). & $\frac{1}{2}$ &  $\frac{1}{2} + h(d-\frac{1}{2}) $ & $d$ 
\end{tabular}\\
where $d > 1/2$ and $h$ ($0 \geq h \geq 1$) acts to specify the dominance of the $B$
allele's effect on female meiotic drive through the males genotype. We
set up a set of recursions using a recessive sex independent cost to
the driver, similar to those in eqn.
\ref{eqn.A1}-\ref{eqn.A1}, but now featuring these male genotype
transmission probabilites. In Supplementary Figure \ref{Effect_of_dominance} we show the invasion
criteria for both the A and the B allele when they are rare in the
population. We find that there is a narrow parameter range the B allele
can invade but not fix if it is dominant (i.e. $h>0$) and the cost in homozygotes
is not too great. This reflects the fact that when the allele is rare
it can increase in frequency due to fertilization of heterozygous eggs
by sperm from heterozygote males, where the sperm do not themselves
carry the allele. But as the B allele becomes common it will frequently be
present in the homozygous state in males, and so will be unable to
avoid the production of unfit homozygote zygotes leaving it trapped in
the population. \gc{Perhaps we should show the equil. freq. in the
  fig, as I'm guessing it might be rare}

>>>>

%\title{Why does female meiosis allow the opportunity for exploitation by selfish sperm? BETTER TITLE?}
\title{
%Co-operation between the sexes trumps selfish control of female  meiotic drive. 
%Or 
Sperm do not evolve to collaborate in female meiotic drive. 
%Or Why does female meiosis allow the opportunity for exploitation by selfish sperm? 
}

@@

Despite the apparent unity of the organism, alleles can
gain an evolutionary advantage at a cost to  individual fitness
\citep{Burt2006}, often by exploiting meiosis and gametogenesis.
Because only one of the four products of female meiosis is transmitted to the egg, female meiosis is particularly vulnerable to such exploitation \citep{Sandler1957,Pardo-ManuelDeVillena2001a}. 
An allele that biases female meiosis in its favor (i.e. a meiotic driver), may increase in frequency even if this driver entails a pleiotropic fitness cost \citep{Prout1973}, generating a genetic conflict between the success of the driver and organismal fitness.
Meiotic drivers observed in nature (in both plants
\citep{Buckler1999,Fishman2005,Fishman2008}, and animals
\citep{Agulnik1990,Wu2005,Pardo-ManuelDeVillena2001c}) highlight this
conflict -- the selfish benefits of drive and the associated
pleiotropic fitness costs sustain a balanced polymorphism
\citep{Prout1973}, 
and often generate ongoing evolutionary escalations of drive suppressors and enhancers \citep{Dawe1996,Fishman2008}. 
The threat of meiotic drive to organismal fitness is potentially so
severe that many basic properties of
meiosis and o\"{o}genesis, including the initial genome doubling in
meiosis I \citep{Haig1991}, arrested female meiosis
\citep{Mira1998}, centromere machinery \citep{Malik2002a,Malik2009,Malik2009}, and sex differences in the recombination rate \citep{Haig2010,Brandvain2012} 
have perhaps evolved to enforce fairness by disrupting meiotic drive \citep{Rice2013}.

@@

It is certainly not the case that animals are mechanistically incapable of circumventing this requirement.
The stage at which female meiosis requires fertilization varies widely across taxa, and in a few animal taxa, 
	female meiosis is complete upon ovulation (that is fertilization is not required for female meiosis \citep[see table 1 of ref 
][]{Masui_book}).

It is also not the case that sperm is mechanistically incapable of influencing female meiosis.
In \emph{C. elegans} premature deployment of the aster (a vital component of mitotic
machinery) provided by the sperm can disrupts MII meiotic segregation
in the egg, leading to a triploid zygote \citep{McNally2012}. 
Thus, there is direct mechanistic evidence for that sperm can influence female meiosis.
 Additionally, genetic evidence from the In and Om  drive system sin mice suggests that the 
 transmission of alternate alleles in heterozygous females can depend on sperm haplotype.  
 Both of these female drive systems operate by distorting the second meiotic division, 
 and in both systems the outcome of female meiosis depends the genotype of the fertilizing sperm \citep{Agulnik1993,Wu2005}.  
While such genetic evidence is suggestive, we note the difficulty in
ruling out the alternative hypothesis of early selection on zygotes \citep[pages
52-54][]{Burt2006}.

@@

In the standard single-locus, biallelic model of female meiotic drive, the driving allele is transmitted to the egg  in heterozygotes with probability  $d > \frac{1}{2}$, regardless of sperm genotype. 
To depict a case of a self-promoting meiotic
driver,  we modify this standard model such that the driver is only effective when fertilized by a sperm carrying that allele (see Figure
\ref{Eggsperm_2_allele_cartoon}) \footnote{the timing of fertilization relative to female meiosis places another constraint on $d$, for example, if fertilization (and therefore, sperm dependent drive) takes place at MII (as in mammals),
	female drive requires an uneven number of crossovers between the centromere and the drive locus, 
	so $d$ is bounded to be $<0.75$? \citep[see ][ for discussion]{Buckler1999}. Could also put this in
        caption of fig.}.
We then identify the conditions under which this self propagating driver can spread, and evaluate whether a driver of this form could generate a sustained conflict favoring the evolution of suppressors. 
\yb{EQUATIONS A1-A4}

We first consider a driving allele neutral in heterozygotes, 
	but deleterious in homozygotes (whose relative fitness is
        $1-s$). This model is laid out in eqn. \ref{eqn.A1}-\ref{eqn.A3}.
In this model, a standard meiotic driver can always invade because 
	when rare it occurs predominantly in heterozygotes and it therefore drives without a fitness cost. 
However, a severe homozygous fitness cost ($s>XXd$) prevents  
	drivers from fixing  \citep{Prout1973}, 
	providing an opportunity for the evolution of
	drive suppressors (CITE - citep{XX}), corresponding well to 
	empirical examples of female meiotic drive.

In contrast to a traditional driver, which drives but pays effectively
no fitness cost as a rare deleterious recessive, 
	a self-promoting driver creates homozygotes
        when it drives, and therefore there is a drive-associated fitness cost even when rare. 
The self-promoting driver must overcome this homozygous fitness cost simply to spread when rare.
Therefore with a recessive fitness cost,
	the conditions allowing the invasion of a self-promoting driver
 	are far more restrictive than those for a standard meiotic driver (see Figure \ref{Invasion_space}).

@@

Although the allelic identity of sperm could plausibly influence the outcome of female meiosis, 
	limited gene expression in sperm \citep[e.g.][]{Joseph2004}
	suggest a model where sperm influence female meiosis via expression of the fertilizing male's
	diploid genotype (perhaps due to proteins and RNAs packaged into the sperm).
This adds an additional parameter to the system -- the dominance of the drive enhancer in heterozygous males, 
	when this dominance is high, a narrow sliver of parameter space sustaining a polymorphism emerges (see Figure S\ref{Effect_of_dominance})
 presumably because positive frequency dependence diminishes with the dominance of the male effect driver.

@@

Above, we explored a model were a driver that drove in females when signaled by a complementary allele in sperm.  
We complement this single locus model of an allele with pleiotropic effects in both sexes, 
 with  an alternative model of two linked loci - one female driver and the other a sperm allele which modifies the effect of drive upon fertilization. 
%Here, we examine an alternative scenario (and more realistic) in which
In this model, a female meiotic driver with no sperm-dependency, initially reaches drive-viability equilibrium (with two alleles
$A$ and $B$ are the ancestral non-driver and driver alleles).
Subsequently, a sperm-acting modifier of female meiotic drive arises tightly
        linked to this balanced locus (effectively creating a third
        allele/haplotype at this locus).
Assuming tight linkage between female driver and sperm modifier is consistent with the nature of well characterized meiotic drive systems 
	which are often maintained as polymorphic inversions with numerous linked modifiers (CITE).

@@

%%%sperm-acting facilitator
We first consider a sperm-acting facilitator of
	female meiotic drive tightly linked to the $B$ background, 
	generating a third allele $B^{+}$ (i.e. the non-recombining haplotype of a
        driver allele and sperm-acting drive modifier). 
This new allele acts in sperm to increase the effectiveness of both
	the $B$ and  $B^{+}$ drive in $AB$ and $AB^{+}$ heterozygotes, see Figure \ref{Eggsperm_3_allele_cartoon} 
	for a simple schematic of this drive model.  
Naively, the $B^{+}$ may spread by capitalizing on the additional drive it causes; however,  
	this is not the case for a few simple reasons. 
First, the sperm-acting drive facilitator (the novel $B^{+}$ haplotype) 
	arises when the ancestral driver ($B$) is at drive-selection balance, 
	and therefore immediately suffers a homozygous fitness cost.  
Worse yet, a novel $B^{+}$ haplotype most often helps 
	the $B$  allele drive ($B^+$ sperm meeting $AB$ eggs), because $B$ is initially more common than $B^{+}$. 
As $B^{+}$ facilitates $B$ these interactions frequently form 
	$BB^{+}$ heterozygotes that suffer the same fitness costs as $BB$. 
Therefore, sperm-acting drive facilitator alleles experience a profound disadvantage 
	in this scenario, even more so than under the previous two allele model. 
We have found no parameter range of this
	three allele system that allows the sperm-acting drive facilitator $B^{+}$ to
	invade the population. \yb{MATHS?}

@@

Sexual reproduction is a high-stakes event that determines what gets transmitted to the next generation.  
As a consequence of this intense competition, alleles that gain a transmission advantage during reproduction 
	can succeed evolutionarily, even if they lower the 
        fitness of the population, 
	the mother whose offspring they will inhabit, 
	or the organism in which they reside. 
This generates numerous conflicts  \cite{Burt2006} including sexual conflicts between mates \cite{Arnqvist2005}, 
	and conflicts between alleles that are overtransmitted in meiosis and the organisms they injure while doing so. 
Such conflicts and their resolution likely play a major role in the structure and evolution of many basic biological processes \citep{Rice2013}.

It seems that allowing sperm to influence the outcome of female meiosis would generate a confluence of these potential conflicts -- 
	sperm could actually assist an allele that distorts female meiosis.
However, this is not the case.
We find that an allele which acts through sperm to distort female meiosis in its favor %YB [too awkward / unclear?] 
	can rarely spread through a population if it bears any cost. 
Additionally, when this self-promoting driver can spread, it can only rarely 
	be maintained as a protected polymorphism, and due to its positive frequency dependence 
	it spends very little time at intermediate frequency.
As such, this type of exploitation cannot generate a sustained genetic conflict.
It is therefore unlikely
	that female oogenesis and meiosis will evolve to prevent their effect.  
This likely explains why females can delay the completion of meiosis until after fertilization 
	without risking exploitation by collaborations by female
        drivers and sperm alleles.
%Male modification of recombination rates. \yb{trying to think through this \dots{}}

Why is it that an allele that biases female meiosis in its favor can generate a genetic conflict, but an allele in sperm that assists this female driver cannot? 
So long as the transmission advantage of female meiotic drive outweigh the organismal fitness cost to heterozygotes the female driver can spread when rare, and it increases in 		
	frequency until the fitness cost experienced when homozygous balances the transmission advantage.
By contrast a sperm promotor of female drive is only effective when matched with a heterozgyote female -- meaning that when rare this allele rarely enhances female drive. 
Even worse, when it does so it will preferentially find itself in a low fitness drive homozygote. 
Not only are drive-promoting sperm alleles unable to create a sustained genetic conflict, 
	but alleles in sperm with the opposite effect - that is those that prevent their own drive through female meiosis do maintain a polymorphisms and 
	provide evolution  with time and opportunity to further minimize drive.
This is because such drive suppressing alleles reduced their chances
of forming low fitness homozygotes. 
In fact, an allele in sperm which causes the alternate allele to drive can be favored (FIG), a counterintuitive result 
	consistent with the tenuous evidence that the non-driving
        $ln^-$ allele is over transmitted through female meiosis in $ln^-/ln^+$ heterozygote when fertilized by $ln^+$
        sperm (CITE cite{agulnick, Pomiankowski1993}).   
More generally, alleles that act through sperm to
 reduce the opportunity of female meiotic drive regardless of linkage or phase.

%More generally both linked and unlinked alleles may often spread that act through sperm to
%reduce the opportunity of female meiotic drive in order to disable
%segregating drive systems, by for example reducing the mechanistic
%asymmetries that female drivers exploit.

@@

%Gametogenesis and meiosis are common times for genetic conflict. 
%This is a consequence of the fact that as the set of alleles that consitute
%	an individual are in the process of parting ways, 
%	with only a subset of them being assured transmission.
%In constrast, the perspective of the fertilizing sperm's genome is different, it has to live with the consequences of any distortion it causes 
%to female meiosis through the entire next life cycle of the organism. 
%It is therefore in the interest of the fertilizing sperm's genome to
%make a high fitness zygote even if that means alleles in sperm
%disrupting other copies of themselves from gaining a 
%transmission advantage through female meiosis.
  \appendix
\section*{Appendix}
In our first scenerio allele $A$ is the ancestral allele and allele
$B$ is our sperm dependent drive allele.
The sperm dependence of the drive is a form of assortative
transmission (mating), and so we have to keep track of the genotype frequencies.
Our genotype frequencies at birth are $x_{AA},~x_{AB},$ and
$x_{BB}$. These three genotypes have fitnesses $w_{AA},~w_{AB},$ and
$w_{BB}$, which are identical between males and females. Because 
fitnesses are equivalent in both sexes, genotypic frequencies
are identical in males and females, but in the following
equations they are marked with $\Mars$ and $\Venus$ respectively to
allow the reader to more easily follow the logic. 
The frequency of the three genotypes at birth in the next generation is:
\begin{equation}
x'_{BB} =\frac{1}{\bar{w} } \Big( w_{BB} x_{BB}^{\Venus} (w_{BB} x_{BB}^{\Mars} +\frac{1}{2}
  w_{AB} x_{AB}^{\Mars} ) + w_{AB} x_{AB}^{\Venus} (d w_{BB}
  x_{BB}^{\Mars} +\frac{1}{2} d
w_{AB} x_{AB}^{\Mars} ) \Big),  \label{eqn.A1}
\end{equation}
\begin{equation}
x'_{AA} =\frac{1}{\bar{w} } \Big(  w_{AA} x_{AA}^{\Venus} (w_{AA} x_{AA}^{\Mars} + \frac{1}{2}
   w_{AB} x_{AB}^{\Mars}) + w_{AB} x_{AB}^{\Venus}(\frac{1}{2} w_{AA} x_{AA}^{\Mars}+ \frac{1}{2} 
  w_{AB} x_{AB}^{\Mars}) \Big),  \label{eqn.A2}
\end{equation}
and
\begin{eqnarray}
x'_{AB} = & \frac{1}{\bar{w} } \Big( w_{AA} x_{AA}^{\Venus}  ( w_{BB}  x_{BB}^{\Mars} + \frac{1}{2}
  w_{AB}  x_{AB}^{\Mars} ) +w_{BB} x_{BB}^{\Venus} ( w_{AA}  x_{AA}^{\Mars} + \frac{1}{2}
  w_{AB}  x_{AB}^{\Mars} )  \nonumber \\
& + x_{AB}^{\Venus}(\frac{1}{2}  w_{AA}  x_{AA}^{\Mars}+\frac{1}{2} (1-d)
 w_{AB}  x_{AB}^{\Mars} + \frac{1}{4}  w_{AB}
 x_{AB}^{\Mars} + (1-d)  w_{BB}  x_{BB}^{\Mars} ) \Big) \label{eqn.A3}
\end{eqnarray}
where the mean fitness of the population is
\begin{equation}
\bar{w} = w_{AA} x_{AA}+w_{AB}  x_{AB} +w_{BB}  x_{BB}.
\end{equation}
We can then iterate this recursion over generations, for
example to produce the frequency trajectories shown in the inset
Figure \ref{Invasion_space} by plotting $x_{BB}+ \frac{1}{2}x_{AB}$
over time. We can also use this to assess the
invasion criteria for the allele by starting allele B at very low
frequency, and seeing if it increases across the generations.

\paragraph{cost of driver in heterozygotes}
We also explored a range of other costs to the drive alleles other than a
simple recessive cost, for example an additive cost. We modified the recursions given in eqn
\ref{eqn.A1}-\ref{eqn.A3} to feature an additive cost $s$ for each
copy of the $B$ allele. A none recessive cost to the drive allele can
cause bistability of the drive allele. The B allele is unable to enter
the population when rare, this results from the fact that
when the allele is rare ($x_B \ll 1$) sperm carrying it rarely
encounter heterozygous eggs ($\sim x^2$) but the allele still pays a
cost in heterozygous individuals.  By contrast, when the B allele is common ($x_B
\gg 0$) encounters between heterozygous eggs and $B$ allele sperm
become more common allowing drive to occur this can allow the $B$
allele to spread when common. In Supplementary Figure
\ref{bistable_additive} we show the unstable equilibrium frequency
above which the $B$ allele can invade. This equilibrium frequency was found by a simple
parameter search.  \gc{wonder if a bit of work could find this}

\paragraph{Dominance} In the supplement we also explore a model where
the genotype of the male who produced the fertilizing sperm affects female drive,
instead of just the allelic type of the sperm.
 We assume a heterozygous female will transmit the $B$ allele 
  with probabilities  $\frac{1}{2}$,  $\frac{1}{2} + h(d-\frac{1}{2}) $ or $d$, 
 if she mated with a n $AA$, $AB$, or $BB$ male,  respectively.
 Above, $d > 1/2$ and $h$ ($0 \geq h \geq 1$) acts to specify the dominance of the $B$
%allele's effect on female meiotic drive through the males genotype.
%YB TRY_1 We assume that in a mating between a heterozygous female and an $AA$, $AB$, or $BB$ male, the female will transmit the $B$ allele 
% with probabilities  $\frac{1}{2}$,  $\frac{1}{2} + h(d-\frac{1}{2}) $, and $d$, respectively.
  %Here $d > 1/2$ and $h$ ($0 \geq h \geq 1$) acts to specify the dominance of the $B$
%allele's effect on female meiotic drive through the males genotype.
% GC ORIGINAL  We assume that if a
%heterozygous egg ($AB$) is fertilized by a sperm with the following
%male ($\Mars$) genotypes the $B$ allele will be transmitted through
%female meiosis with the following probabilities \\
%\begin{tabular}{cccc}
%$\Mars$ geno. & AA & AB  & BB\\
%P(B). & $\frac{1}{2}$ &  $\frac{1}{2} + h(d-\frac{1}{2}) $ & $d$ 
%\end{tabular}\\
%where $d > 1/2$ and $h$ ($0 \geq h \geq 1$) acts to specify the dominance of the $B$
%allele's effect on female meiotic drive through the males genotype. 
We set up a set of recursions using a recessive sex independent cost to
the driver, similar to those in eqn.
\ref{eqn.A1}-\ref{eqn.A1}, but now featuring these male genotype
transmission probabilites. In Supplementary Figure \ref{Effect_of_dominance} we show the invasion
criteria for both the A and B alleles when they are rare. 
We find a narrow parameter range where B 
can invade but not fix.
This corresponds to B being less than fully recessive (i.e. $h>0$) and to a modest cost in homozygotes. 
This combination of parameters reflects the fact that when the allele is rare, 
it can increase in frequency due to fertilization of heterozygous eggs
by sperm from heterozygote males, where the sperm do not themselves
carry the allele. But as the B allele becomes common it will frequently be
present in the homozygous state in males, and so will be unable to
avoid the production of unfit homozygote zygotes leaving it trapped in
the population. \gc{Perhaps we should show the equil. freq. in the
  fig, as I'm guessing it might be rare}
